Detecting Stylistic Variation in Pop Production using the Spotify API

### 2023-11-13 outputs/paper/paper.qmd
PDF render

---
title: "Detecting Stylistic Variation in Pop Production using the Spotify API"
author: 
  - Michaela Drouillard
thanks: "Code and data are available at: https://github.com/michaeladrouillard/spotify_pop."
date: today
abstract: "This paper explores whether or not we can predict whether a song was produced by Jack Antonoff using the audio features provided by Spotify's API. We find that in general our model has an accuracy score of 90%. (Precision and recall) We interestingly found that danceability, a metric trained on human-tagged data, was a key predictor. Our results have implications for the understanding of culture, how the combination human subjectivity and machine listening can come to create an approximate for style."
format: pdf
number-sections: true
bibliography: references.bib
---

@@

```{r}
#| label: tbl-producercounts
#| tbl-cap: Counts of Tracks Per Producer in Original Dataset
#| echo: false
#| warning: false

@@

Macdonald suggests combining energy and valence to create quadrant, which usually works "fairly well" to describe music. High energy and high valence could be generally happy, cheerful, upbeat. Low energy, low valence could be sad or downbeat. High energy, low valence is sort of angry. Low energy, high valence is serene or calming.

```{r}
#| label: fig-quadvalence
#| fig-cap: Valence and Energy Quadrant
#| echo: false
#| warning: false
av<- read_csv(here::here("inputs/data/antonoff_viz.csv"))

@@

ggplot(av, aes(x = energy, y = valence)) +
  geom_point(aes(color = as.factor(is_antonoff)), alpha = 0.3) +
  geom_vline(xintercept = mean(df$energy), color = "gray") +
  geom_hline(yintercept = mean(df$valence), color = "gray") +
  theme_minimal()
```

>>>>

---
title: "Detecting Stylistic Variation in Pop Production using the Spotify API"
author: 
  - Michaela Drouillard
thanks: "Code and data are available at: https://github.com/michaeladrouillard/spotify_pop."
date: today
abstract: "This paper explores whether or not we can predict whether a song was produced by Jack Antonoff using the audio features provided by Spotify's API. We find that in general our logistic regression model has an accuracy score of 76%, a precision score of 39%, and a recall of 75.61%. Our random forest has an accuracy of 61%, and a No Information Rate of 43%. We interestingly found that danceability, a metric trained on human-tagged data, was a key predictor. Our results have implications for computational explorations of culture, how the combination human subjectivity and machine listening can come to create an approximate for style."
format: pdf
number-sections: true
bibliography: references.bib
---

@@

```{r}
#| label: tbl-producercounts
#| tbl-cap: Counts of Tracks Per Producer in Original Dataset
#| echo: false
#| warning: false
lr_data<- read_csv(here::here("inputs/data/lr_data.csv"))

@@

Macdonald suggests combining energy and valence to create quadrant, which usually works "fairly well" to describe music (see @fig-quadvalence). High energy and high valence could be generally happy, cheerful, upbeat. Low energy, low valence could be sad or downbeat. High energy, low valence is sort of angry. Low energy, high valence is serene or calming.

```{r}
#| label: fig-quadvalence
#| fig-cap: Valence and Energy Quadrant
#| echo: false
#| warning: false
av <- read.csv(here::here("inputs/data/antonoff_viz.csv"))

@@

ggplot(av, aes(x = energy, y = valence)) +
  geom_point(aes(color = as.factor(is_antonoff)), alpha = 0.3) +
  geom_vline(xintercept = mean(av$energy), color = "gray") +
  geom_hline(yintercept = mean(av$valence), color = "gray") +
  theme_minimal()
```
